---
title: "ibd_analysis"
author: "Angela Jiang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(sjPlot)

```

## R Markdown

### Read in files

```{r read}
ibd_mtx = read.csv("mtx_all.tsv", sep="\t")
ibd_mgx = read.csv("diagnosis_new_mgx.tsv", sep="\t")
```

### Process data

Removing duplicates, remove samples with <1mil reads, normalizing by calculating count per million

```{r process}
ibd_mtx <- unique(ibd_mtx[,c(2,3,4,5)])

ibd_mgx <- unique(ibd_mgx[,c(2,3,4,5)])

ibd_mtx_processed <- ibd_mtx %>%
  filter(total_reads >= 5000000) %>%
  mutate(cpm = total_mapped / total_reads * 1000000)

ibd_mgx_processed <- ibd_mgx %>%
  filter(total_reads >= 5000000) %>%
  mutate(cpm = total_mapped / total_reads * 1000000)

ibd_mtx_processed[ibd_mtx_processed == 0] <- 0.0001

ibd_mgx_processed[ibd_mgx_processed == 0] <- 0.0001
```

### Normality test
```{r normality, echo=FALSE}
shapiro.test(ibd_mgx_processed$cpm)

shapiro.test(ibd_mtx_processed$cpm)
```

### MGX plot

Plotting MGX data

```{r mgx_plot, echo=FALSE}
p <- ggplot(ibd_mgx_processed, aes(x=diagnosis, y=cpm)) + 
  geom_boxplot() +
  ylab("log(cpm)") +
  scale_y_log10() +
  theme_classic()

p

save_plot("ibd_mgx.svg", fig = p, width=10, height=10)
```

### MTX plot

Plotting MTX data

```{r mtx_plot, echo=FALSE}
t <- ggplot(ibd_mtx_processed, aes(x=diagnosis, y=cpm)) + 
  geom_boxplot() +
  ylab("log(cpm)") +
  scale_y_log10() +
  theme_classic()

t

save_plot("ibd_mtx.svg", fig = t, width=10, height=10)
```
### Filter data into comparison groups

```{r mtx_stat, echo=FALSE}
#MTX
UCvsnonIBD_mtx <- ibd_mtx_processed %>%
  filter(diagnosis == "UC" | diagnosis == "nonIBD")

CDvsnonIBD_mtx <- ibd_mtx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "nonIBD")

CDvsnonUC_mtx <- ibd_mtx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "UC")

#MGX
UCvsnonIBD_mgx <- ibd_mgx_processed %>%
  filter(diagnosis == "UC" | diagnosis == "nonIBD")

CDvsnonIBD_mgx <- ibd_mgx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "nonIBD")

CDvsnonUC_mgx <- ibd_mgx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "UC")
```

## MTX stats
### UC vs non-IBD MTX wilcox test

```{r mtx_uc_nonibd, echo=FALSE}
wilcox.test(total_mapped~ diagnosis, 
            data = UCvsnonIBD_mtx,
            exact = FALSE)
```

### CD vs non-IBD MTX wilcox test

```{r mtx_cd_nonibd, echo=FALSE}
wilcox.test(total_mapped~ diagnosis, 
            data = CDvsnonIBD_mtx,
            exact = FALSE)
```

### CD vs UC MTX wilcox test

```{r mtx_cd_uc, echo=FALSE}
wilcox.test(total_mapped~ diagnosis, 
            data = CDvsnonUC_mtx,
            exact = FALSE)
```

## MGX stats
### UC vs non-IBD MGX wilcox test

```{r mgx_uc_nonibd, echo=FALSE}
wilcox.test(total_mapped~ diagnosis, 
            data = UCvsnonIBD_mgx,
            exact = FALSE)
```

### CD vs non-IBD MGX wilcox test

```{r mgx_cd_nonibd, echo=FALSE}
wilcox.test(total_mapped~ diagnosis, 
            data = CDvsnonIBD_mgx,
            exact = FALSE)
```

### CD vs UC MGX wilcox test

```{r mgx_cd_uc, echo=FALSE}
wilcox.test(total_mapped~ diagnosis, 
            data = CDvsnonUC_mgx,
            exact = FALSE)
```

