---
title: "ibd_analysis"
author: "Angela Jiang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(sjPlot)

```

## R Markdown

### Read in files

```{r read}
ibd_mtx = read.csv("Tables/mtx_total_diagnosis.tsv", sep="\t")
ibd_mgx = read.csv("Tables/mgx_diagnosis.tsv", sep="\t")
```

### Process data

Removing duplicates, remove samples with <1mil reads, normalizing by calculating count per million

```{r process}
ibd_mtx <- unique(ibd_mtx[,c(2,3,4,5)])

ibd_mgx <- unique(ibd_mgx[,c(2,3,4,5)])

ibd_mtx_processed <- ibd_mtx %>%
  filter(total_reads >= 5000000) %>%
  mutate(cpm = total_mapped / total_reads * 1000000)

ibd_mgx_processed <- ibd_mgx %>%
  filter(total_reads >= 5000000) %>%
  mutate(cpm = total_mapped / total_reads * 1000000)

ibd_mtx_processed[ibd_mtx_processed == 0] <- 0.0001

ibd_mgx_processed[ibd_mgx_processed == 0] <- 0.0001



```

### Normality test
```{r normality, echo=FALSE}
shapiro.test(ibd_mgx_processed$cpm)

shapiro.test(ibd_mtx_processed$cpm)
```

### MGX plot

Plotting MGX data

```{r mgx_plot, echo=FALSE}
p <- ggplot(ibd_mgx_processed, aes(x=diagnosis, y=cpm)) + 
  geom_point() +
  geom_jitter(position = position_jitter(width = 0.1, height = 0.1))+
  ylab("log(cpm)") +
  #scale_y_log10() +
  theme_classic()

p <- ggplot(ibd_mgx_processed, aes(x=diagnosis, y=cpm, fill=diagnosis)) +
    geom_violin(width=1) +
    geom_boxplot(width=0.1, color="grey", alpha=0.2, outlier.shape = NA) +
  scale_fill_brewer(palette = "Dark2") +
    #scale_fill_viridis(discrete = TRUE) +
    ggtitle("MGX") +
    scale_y_log10() +
    xlab("Diagnosis") +
   ylab("log(cpm)") +
  theme_classic()

p

#save_plot("ibd_mgx.svg", fig = p, width=10, height=10)
ggsave(file="ibd_mgx.svg", plot=p, width=6, height=6)


```

### MTX plot

Plotting MTX data

```{r mtx_plot, echo=FALSE}
t <- ggplot(ibd_mtx_processed, aes(x=diagnosis, y=cpm)) + 
  geom_boxplot(fill="gray") +
  ylab("log(cpm)") +
  scale_y_log10() +
  theme_classic()

t <- ggplot(ibd_mtx_processed, aes(x=diagnosis, y=cpm, fill=diagnosis)) +
    geom_violin(width=1) +
    geom_boxplot(width=0.1, color="grey", alpha=0.2, outlier.shape = NA) +
  scale_fill_brewer(palette = "Dark2") +
    #scale_fill_viridis(discrete = TRUE) +
    ggtitle("MTX") +
    scale_y_log10() +
    xlab("Diagnosis") +
    ylab("cpm") +
  theme_classic()

t

ggsave(file="ibd_mtx.svg", plot=t, width=6, height=6)
```
### Filter data into comparison groups

```{r mtx_stat, echo=FALSE}
#MTX
UCvsnonIBD_mtx <- ibd_mtx_processed %>%
  filter(diagnosis == "UC" | diagnosis == "nonIBD")

CDvsnonIBD_mtx <- ibd_mtx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "nonIBD")

CDvsnonUC_mtx <- ibd_mtx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "UC")

#MGX
UCvsnonIBD_mgx <- ibd_mgx_processed %>%
  filter(diagnosis == "UC" | diagnosis == "nonIBD")

CDvsnonIBD_mgx <- ibd_mgx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "nonIBD")

CDvsnonUC_mgx <- ibd_mgx_processed %>%
  filter(diagnosis == "CD" | diagnosis == "UC")
```

## MTX stats
### UC vs non-IBD MTX wilcox test

```{r mtx_uc_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis, 
            data = UCvsnonIBD_mtx,
            exact = FALSE)
```

### CD vs non-IBD MTX wilcox test

```{r mtx_cd_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis, 
            data = CDvsnonIBD_mtx,
            exact = FALSE)
```

### CD vs UC MTX wilcox test

```{r mtx_cd_uc, echo=FALSE}
wilcox.test(cpm~ diagnosis, 
            data = CDvsnonUC_mtx,
            exact = FALSE)
```

## MGX stats
### UC vs non-IBD MGX wilcox test

```{r mgx_uc_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis, 
            data = UCvsnonIBD_mgx,
            exact = FALSE)
```

### CD vs non-IBD MGX wilcox test

```{r mgx_cd_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis, 
            data = CDvsnonIBD_mgx,
            exact = FALSE)
```

### CD vs UC MGX wilcox test

```{r mgx_cd_uc, echo=FALSE}
wilcox.test(cpm~ diagnosis, 
            data = CDvsnonUC_mgx,
            exact = FALSE)
```


```{r garL}
ibd_mtx_garl = read.csv("Tables/mtx_gudl.tsv", sep="\t")
ibd_mgx_garl = read.csv("Tables/gudl_diagnosis_mgx.tsv", sep="\t")


```

```{r garL}
ibd_mtx_garl = merge(ibd_mtx_garl,ibd_mtx, by="External.ID")

ibd_mgx_garl = merge(ibd_mgx_garl,ibd_mgx, by="External.ID")


ibd_mtx_garl <- ibd_mtx_garl %>%
  filter(total_reads >= 5000000) %>%
  mutate(cpm = garl_reads / total_reads * 1000000)

ibd_mgx_garl <- ibd_mgx_garl %>%
  filter(total_reads >= 5000000) %>%
  mutate(cpm = garl_reads / total_reads * 1000000)

ibd_mtx_garl[ibd_mtx_garl == 0] <- 0.0001

ibd_mgx_garl[ibd_mgx_garl == 0] <- 0.0001

```

``` {r graph_mtx_garl}
p <- ggplot(ibd_mtx_garl, aes(x=diagnosis.x, y=cpm, fill=diagnosis.x)) +
    geom_violin(width=1) +
    geom_boxplot(width=0.1, color="grey", alpha=0.2, outlier.shape = NA) +
  scale_fill_brewer(palette = "Dark2") +
    #scale_fill_viridis(discrete = TRUE) +
    ggtitle("MTX") +
    scale_y_log10() +
    xlab("Diagnosis") +
    ylab("cpm") +
  theme_classic()

p

p2 <- ggplot(ibd_mgx_garl, aes(x=diagnosis.x, y=cpm, fill=diagnosis.x)) +
    geom_violin(width=1) +
    geom_boxplot(width=0.1, color="grey", alpha=0.2, outlier.shape = NA) +
  scale_fill_brewer(palette = "Dark2") +
    #scale_fill_viridis(discrete = TRUE) +
    ggtitle("MTX") +
    scale_y_log10() +
    xlab("Diagnosis") +
    ylab("cpm") +
  theme_classic()

p2

ggsave(file="ibd_mgx_garl.svg", plot=p2, width=6, height=6)
ggsave(file="ibd_mtx_garl.svg", plot=p, width=6, height=6)


```

### Filter data into comparison groups

```{r mtx_stat, echo=FALSE}
#MTX
UCvsnonIBD_mtx_garl <- ibd_mtx_garl %>%
  filter(diagnosis.x == "UC" | diagnosis.x == "nonIBD")

CDvsnonIBD_mtx_garl <- ibd_mtx_garl %>%
  filter(diagnosis.x == "CD" | diagnosis.x == "nonIBD")

CDvsnonUC_mtx_garl <- ibd_mtx_garl %>%
  filter(diagnosis.x == "CD" | diagnosis.x == "UC")

#MGX
UCvsnonIBD_mgx_garl <- ibd_mgx_garl %>%
  filter(diagnosis.x == "UC" | diagnosis.x == "nonIBD")

CDvsnonIBD_mgx_garl <- ibd_mgx_garl %>%
  filter(diagnosis.x == "CD" | diagnosis.x == "nonIBD")

CDvsnonUC_mgx_garl <- ibd_mgx_garl %>%
  filter(diagnosis.x == "CD" | diagnosis.x == "UC")
```

## MTX stats
### UC vs non-IBD MTX wilcox test

```{r mtx_uc_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis.x, 
            data = UCvsnonIBD_mtx_garl,
            exact = FALSE)
```

### CD vs non-IBD MTX wilcox test

```{r mtx_cd_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis.x, 
            data = CDvsnonIBD_mtx_garl,
            exact = FALSE)
```

### CD vs UC MTX wilcox test

```{r mtx_cd_uc, echo=FALSE}
wilcox.test(cpm~ diagnosis.x, 
            data = CDvsnonUC_mtx_garl,
            exact = FALSE)
```

## MGX stats
### UC vs non-IBD MGX wilcox test

```{r mgx_uc_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis.x, 
            data = UCvsnonIBD_mgx_garl,
            exact = FALSE)
```

### CD vs non-IBD MGX wilcox test

```{r mgx_cd_nonibd, echo=FALSE}
wilcox.test(cpm~ diagnosis.x, 
            data = CDvsnonIBD_mgx_garl,
            exact = FALSE)
```

### CD vs UC MGX wilcox test

```{r mgx_cd_uc, echo=FALSE}
wilcox.test(cpm~ diagnosis.x, 
            data = CDvsnonUC_mgx_garl,
            exact = FALSE)
```
